### 2014-09-11 paper/paper.Rmd
fixed code error, hopefully

```{r shortagesbars, fig.width=10, fig.height=12, warning=FALSE, message=FALSE, echo=FALSE}
df$SC14Q02 <- as.character(df$SC14Q02)
df$SC14Q07 <- as.character(df$SC14Q07)
df$SC14Q10 <- as.character(df$SC14Q10)
df$SC14Q02[df$SC14Q02=="A lot"] <- "To some extent"
df$SC14Q07[df$SC14Q07=="A lot"] <- "To some extent"
df$SC14Q10[df$SC14Q10=="A lot"] <- "To some extent"
df$SC14Q02 <- factor(df$SC14Q02, labels=c(0,1,2))
df$SC14Q07 <- factor(df$SC14Q07, labels=c(0,1,2))
df$SC14Q10 <- factor(df$SC14Q10, labels=c(0,1,2))
df.mshort <- df[!is.na(df$SC14Q02),] %>% group_by(name, SC14Q02) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q02, mmath2, data = df.mshort, facets = ~name, 
#      xlab="Math teacher shortage (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df.lshort <- df[!is.na(df$SC14Q10),] %>% group_by(name, SC14Q10) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q10, mmath2, data = df.lshort, facets = ~name, 
#      xlab="Library shortages (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df2 <- summarise(group_by(df[!is.na(df$SC14Q02),], name, SC14Q02), teachers = mean(mmath, na.rm=T))
colnames(df2)[2] <- "levels"
df3 <- summarise(group_by(df[!is.na(df$SC14Q07),], name, SC14Q07), computers = mean(mmath, na.rm=T))
colnames(df3)[2] <- "levels"
df4 <- summarise(group_by(df[!is.na(df$SC14Q10),], name, SC14Q10), libmaterials = mean(mmath, na.rm=T))
colnames(df4)[2] <- "levels"
df.all <- merge(df2, df3)
df.all <- merge(df.all, df4)
df.m <- melt(df.all, id=c("name", "levels"))
colnames(df.m)[4] <- "math"
dqplot(levels, math, data = df.m, facets = ~name, 
      xlab="Shortages (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
      ylab="Math Score", geom=c("point","line"), group=variable, colour=variable) +
      scale_colour_manual("Type", values=c("teachers"="#1B9E77", "computers"="#D95F02", "libmaterials"="#7570B3")) +
      theme(legend.position="bottom") 
```

@@

For each country we have provided loess-smoothed curves of the relationship between the mathematics score and the time spent studying outside of school. The most prominent feature that we observe is the decreasing benefit of studying. For every country there is evidence of a substantial positive relationship for the students who spend less time studying. However, those students who continue to study for long hours see diminishing returns for their efforts. This pattern is consistent across most countries with a few notable exceptions. Most of the Asian countries still have gains for longer hours studied.

>>>>

```{r shortagesbars, fig.width=10, fig.height=12, warning=FALSE, message=FALSE, echo=FALSE}
df$SC14Q02 <- as.character(df$SC14Q02)
df$SC14Q07 <- as.character(df$SC14Q07)
df$SC14Q10 <- as.character(df$SC14Q10)
df$SC14Q02[df$SC14Q02=="A lot"] <- "To some extent"
df$SC14Q07[df$SC14Q07=="A lot"] <- "To some extent"
df$SC14Q10[df$SC14Q10=="A lot"] <- "To some extent"
df$SC14Q02 <- factor(df$SC14Q02, labels=c(0,1,2))
df$SC14Q07 <- factor(df$SC14Q07, labels=c(0,1,2))
df$SC14Q10 <- factor(df$SC14Q10, labels=c(0,1,2))
df.mshort <- df[!is.na(df$SC14Q02),] %>% group_by(name, SC14Q02) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q02, mmath2, data = df.mshort, facets = ~name, 
#      xlab="Math teacher shortage (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df.lshort <- df[!is.na(df$SC14Q10),] %>% group_by(name, SC14Q10) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q10, mmath2, data = df.lshort, facets = ~name, 
#      xlab="Library shortages (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df2 <- summarise(group_by(df[!is.na(df$SC14Q02),], name, SC14Q02), teachers = mean(mmath, na.rm=T))
colnames(df2)[2] <- "levels"
df3 <- summarise(group_by(df[!is.na(df$SC14Q07),], name, SC14Q07), computers = mean(mmath, na.rm=T))
colnames(df3)[2] <- "levels"
df4 <- summarise(group_by(df[!is.na(df$SC14Q10),], name, SC14Q10), libmaterials = mean(mmath, na.rm=T))
colnames(df4)[2] <- "levels"
df.all <- merge(df2, df3)
df.all <- merge(df.all, df4)
df.m <- melt(df.all, id=c("name", "levels"))
colnames(df.m)[4] <- "math"
qplot(levels, math, data = df.m, facets = ~name, 
      xlab="Shortages (0=None, 1=Very little, 2=To some extent, or A lot)", 
      ylab="Math Score", geom=c("point","line"), group=variable, colour=variable) +
      scale_colour_manual("Type", values=c("teachers"="#1B9E77", "computers"="#D95F02", "libmaterials"="#7570B3")) +
      theme(legend.position="bottom") 
```

@@

For each country we have provided loess-smoothed curves of the relationship between the math score and the time spent studying outside of school. Math scores increase on average the more time spent studying, up to about 10 hours. Beyond that scores stay flat, with the exception being several Asian countries: Vietnam, China, South Korea, Taiwan. For China, the gain tails off around 20 hours of study.
